The language of generalization: Probability, vagueness, and interaction

### 2018-03-01 paper/comgen-paper.Rmd
address half of ellens comments in discussion

It is a remarkable fact that so much is learned from ideas expressed vaguely in words.
Generalizations in language (e.g., "John runs.", "Dogs are friendly.", "Staring at the sun makes you go blind.") are a premier example of how simple statements---statements understood by even the youngest language users---can display complex sensitivities to context.
We have argued that the core meaning of such linguistic expressions is simple, but underspecified. 
To our knowledge, this is the first formal theory of generalizations in language that makes precise quantitative predictions about human behavioral data.

As a formal theory of the language of generalization, our framework aims to unify large swaths of language that are, on their surface, quite distinct from one another: "Birds lay eggs", "Sally smokes", "Herb X makes animals sleepy".
<!-- What these expressions have in common is not surface features but generalization (also referred to as *genericity*). -->
We have formalized a unified, core meaning of generalization in language using the underlying scale of predictive probability or *prevalence*.
We showed that endorsements of all three types of generalizations exhibit the same sensitivity to a background distribution comprised of other categories (Expt. 1, 2a \& 2b, 3).
As well, our framework naturally accounts for a number of logically puzzling category generalizations (Expt. 1) and provides an avenue for top-down moderators on language use, by elucidating the semantic scale as defined by a subjective, predictive probability, not mere frequency (Expt. 2c).
Finally, we have shown that listeners' prior knowledge about the prevalence of the feature in other categories is directly influences their endorsement of generalizations for the same category (Expt. 3).
<!-- In addition to unifying seemingly disparate parts of language, this paper provides a theoretical framework for asking further questions about genericity, which we outline below. -->

In this paper, we proposed a model for interpreting and endorsing generalizations in language by formalizing a simple semantic theory in a communicative model with shared background knowledge.
Our pair of models can be seen as a special case of a Rational Speech Act model, a rational framework of pragmatic language understanding [@Frank2012; @Goodman2016].
The endorsement model, which we use to generate predictions for our three case studies, is a special case of a speaker model who can only produce one of two utterances: the generalization or silence. 
The interpreter model is a "literal listener", who updates its beliefs by averaging over plausibly thresholds given prior knowledge and the simple truth-functional meaning.

<!-- The theory we present here goes beyond previous accounts in a number of useful ways. -->
<!-- The foremost contribution of our theory is that it is accompanied by a formal model, which makes precise quantitative predictions about endorsments. -->
The formal Bayesian model we articulate provides a clean separation of the semantics of a generalization from background, world knowledge.
This formal separation is a key advancement beyond previous accounts, which have either aimed to account for context-sensitivity by positing distinct semantic constructs [e.g., *relative* vs *absolute* generics, described below; @Cohen1999] or positing a semantics that cannot be separated from world knowledge [@Leslie2008].
Our model formalizes a simple meaning for generalizations and investigates the structure of world knowledge that could give rise to the context-sensitivities observed in the data.
That is, we provide a route for precise quantitative influence of conceptual knowledge on generic language by way of shared background knowledge between speakers and listeners.
In the rest of this discussion, we elaborate some key features of the theory that warrant further investigation, sketch out arguments for other known philosophical puzzles in genericity, and situate our theory with respect to extant theories of genericity.

@@

Our theory assumes that listeners and speakers share relevant background knowledge (in the form of a *prevalence prior*) against which generalizations are interpreted.
Given this representation of background knowledge (*prevalence prior*), language interfaces with conceptual knowledge via *prevalence*. 
Previous psychological and philosophical work on generics has looked beyond prevalence and focused on kinds and their conceptual relations to properties [@Leslie2007; @Leslie2008; @Prasada2013].
Indeed, we see generics as describing kinds, but argue that the reason why particular generics are licensed in a context concerns the information conveyed to a listener relative to their background knowledge, formalized with by prevalence prior.
In the right situations, this shared background knowledge can presuppose beliefs about natural kinds and richly structured relations between categories and the properties [a la @Prasada2013].
In other situations, relevant background knowledge may be merely correlational in nature.

@@

The probabilistic communicative model we introduce in this paper assumes shared background 
<!-- Our model of an interpreter, or naive listener, has access to implicit  -->
statistics of the event or property in question, represented by the prior belief distribution over the prevalence $P(h)$.
We constructed prevalence priors for a property, event, or cause by considering *other possible* categories having the property, people doing the action, or causes producing the effect.
Collectively, these other kinds, people, or causes form *comparison classes* against which the referent-category is evaluated.
Thus, the prevalence prior $P(h)$ is actually a conditional distribution constructed with respect some comparison class $C$: $P(h \mid C)$.

@@

Uncertainty about the comparison class adds yet more opportunity for miscommunication, deceit, or just faultless disagreement to the language of generalization.
Even if interlocutors roughly coordinate on a particular comparison class (e.g., agreeing that they are talking *relative to other animals*), they may disagree about the prevalences for other categories in the comparison class.
That is, our model makes the prediction that the acceptance or rejection of controversial statements (e.g., "Muslims are terrorists") depends upon a speaker's beliefs about how often other categories exhibit the property (e.g., neo-nazis committing acts of terrorism).
In Expt. 3, we empirically demonstrated the influence of other categories on endorsement of a generalization.
Understanding how knowledge of the comparison class influences language use is an exciting new area for future research in the language of generalizations.

@@

In this paper, we aimed to balance two competing goals in designing our experimental stimuli: (1) including some familiar, philosophically puzzling examples that have been discussed in the vast literature on generic language and (2) introducing novel stimuli that provide targeted tests of our model vis-a-vis alternative models.
In this section, we discuss how a few other previously-discussed troubling examples of generalizations could be addressed in our communicative framework.
We consider the five following examples:

1. Mary handles the mail from Antarctica.
2. Supreme Court Justices have even social security numbers.
3. Elephants live in Africa and Asia.
4. Books are paperbacks.
5. Women are submissive.

@@

>>>>

It is remarkable that so much is learned from ideas expressed vaguely in words.
Generalizations in language (e.g., "John runs.", "Dogs are friendly.", "Staring at the sun makes you go blind.") are a premier example of how simple statements---statements understood by even the youngest language users---can display complex sensitivities to context.
We have argued that the core meaning of such linguistic expressions can be understood as simple but underspecified. 
To our knowledge, ours is the first formal theory of generalizations in language that makes precise quantitative predictions about human behavioral data.

As a formal theory of the language of generalization, our framework unifies large swaths of language about distinct semantic types: "Robins lay eggs" (categories), "Sally smokes" (events), "Herb X makes animals sleepy" (causes).
Generalization underlies each of these, and we formalize their meaning by appealing to prevalence, which we take to mean *predictive probability* (as opposed to frequency).
We test this theory by exploring its implications for *endorsement*, operationalized as a speaker's decision about whether or not to produce the generalization to naive listener. 
Endorsement, we show, depends in systematic ways on the listener's prior beliefs about the prevalence.
<!-- What these expressions have in common is not surface features but generalization (also referred to as *genericity*). -->
This framework accounts for a number of philosophically puzzling generic statements (Expt. 1), naturally extends to other domains of generalizations (habituals and causals; Expts. 2 \& 3), and provides an avenue for top-down moderators on language use by way of *predictive* probability (Expt. 2c).
We have provided strict tests of our formal model by both measuring and manipulating listeners' background knowledge as well as comparing our model to a number of alternatives.
<!-- In addition to unifying seemingly disparate parts of language, this paper provides a theoretical framework for asking further questions about genericity, which we outline below. -->

<!-- The theory we present here goes beyond previous accounts in a number of useful ways. -->
<!-- The foremost contribution of our theory is that it is accompanied by a formal model, which makes precise quantitative predictions about endorsments. -->
The formal Bayesian model we articulate provides a clean separation of the semantics of a generalization from world knowledge.
This formal separation is a key advancement beyond previous theoretical accounts, which have either aimed to account for context-sensitivity by positing distinct semantic constructs [e.g., *relative* vs *absolute* generics, described below; @Cohen1999] or positing a semantics that cannot be separated from world knowledge [@Leslie2008].
That is, we provide a route for precise quantitative influence of world knowledge on generic language by way of shared background knowledge between speakers and listeners, which at its base takes the form of a prevalence prior.
Our communicative framework also includes a model of a listener who interprets generalization.
Such a quantitative model makes further predictions about domain-general learning from the language of generalization, which we leave for future work. 
In the rest of this discussion, we elaborate how our model relates to other problems in language understanding, sketch out arguments for other known philosophical puzzles in genericity, and situate our theory with respect to extant theories of genericity.

<!-- Our model formalizes a simple meaning for generalizations and investigates the structure of world knowledge that could give rise to the context-sensitivities observed in the data. -->

@@

Our theory assumes that listeners and speakers share relevant background knowledge (in the form of a *prevalence prior*) against which generalizations are interpreted.
Given this representation of background knowledge (*prevalence prior*), language interfaces with conceptual knowledge via *prevalence*. 
Previous psychological and philosophical work on generics has looked beyond prevalence and focused on kinds and their conceptual relations to properties [@Leslie2007; @Leslie2008; @Prasada2013].
Indeed, we see a core communicative function of generics as teaching properties of kinds to naive listeners.
But we argue that the reason why particular generics are licensed in a context concerns the information conveyed to a listener relative to their background knowledge; to interface with a truth-functional semantics, such background knowledge is represented in some capactiy in the form of a prevalence prior.
There is nothing in our theory that rules out shared background knowledge presupposing beliefs about natural kinds and richly structured relations between categories and the properties [a la @Prasada2013].
Such knowledge is not necessary, however: Relevant background knowledge may be merely correlational in nature.

@@

The probabilistic communicative model we introduce in this paper assumes shared background 
<!-- Our model of an interpreter, or naive listener, has access to implicit  -->
statistics of the event or property in question, represented by the prior belief distribution over the prevalence $P(p_{fk})$.
We constructed prevalence priors for a property, event, or cause by considering *other possible* categories having the property, people doing the action, or causes producing the effect.
In Expt. 3, we empirically demonstrated the influence of other categories on endorsement of generalizations.
Collectively, these other kinds, people, or causes form *comparison classes* against which the referent-category is evaluated.
Thus, the prevalence prior $P(p_{fk})$ is actually a conditional distribution constructed with respect some comparison class $C$: $P(p_{fk} \mid C)$.

@@

Uncertainty about the comparison class adds yet more opportunity to the language of generalization for disagreement, miscommunication, or deceit.
Even if interlocutors roughly coordinate on a particular comparison class (e.g., agreeing that they are talking *relative to other animals*), they may fail to coordinate on the prevalences for other categories in the comparison class.
That is, our model makes the prediction that the acceptance or rejection of controversial statements (e.g., "Muslims are terrorists") depends upon a speaker's beliefs about how often other categories exhibit the property (e.g., neo-nazis committing acts of terrorism).
Understanding the contextual influence on the construction of comparison classes should provide insight into failures of communication.
\mht{ellen: TRUE,  I AGREE WITH YOU AND THIS IS PROBABLY OKAY.  BUT IT THINK IT WOULD BE BETTER TO SAY SOMETHING SUBSTANTIVE;  FOCUSSING ON HOW KNOWLEDGE OF THE COMPARISION CLASS SHOULD PROVIDE INSIGHTS INTO FAILURES TO COMMUNICATE SUCCESSFULLY.}

@@

@@

In this paper, we aimed to balance two competing goals in designing our experimental stimuli: (1) including some familiar, philosophically puzzling examples that have been discussed in the vast literature on generic language and (2) introducing novel stimuli that provide targeted tests of our model vis-a-vis alternative models.
We close by discussing how a few other previously-discussed troubling examples of generalizations could be addressed in our communicative framework.
We consider the five following examples:

1. Mary handles the mail from Antarctica (*yet has never had the opportunity*).
2. Supreme Court Justices have even social security numbers.
3. Elephants live in Africa and Asia.
4. Books are paperbacks.
5. Women are submissive.
